Spatiotemporal Multi-Resolution Approximation of the Amari Type Neural Field Model

### 2012-01-22 Multiresolution_framework/ltx/manuscript.tex
added partition unity property of Bsplines

The $m$-th  order cardinal B-spline function is defined by the recurrence relation \cite{Chui1992} 
\begin{equation}
N_{m}\left(r\right) = \left(N_{m-1}\ast N_{1}\right)\left(r\right) = \int_0^{1} N_{m-1}\left( r-r'\right)\,\mathrm{d}r',
\label{SplineConvolutionIntegral}
\end{equation}
where $m>1$, $\ast$ denotes convolution, and $N_1\left(r\right)$ is the characteristic function of the unit interval $\left[ 0,1\right)$, also known as indicator function, i.e.
\begin{equation}
N_{1}\left(r\right)=
\begin{cases}
1 & \text{if $ 0\le r<1$}, \\
0 & \mathrm{elsewhere}.
\end{cases}
\end{equation}
Following this, a B-spline of any order, $N_m(r)$, can be computed using the recursive expression defined by \cite{DeBoor2001}
\begin{equation}\label{eq:MRA-DoBoorFormula}
 N_{m}\left(r\right)=\frac{r}{m-1}N_{m-1}\left(r\right)+\frac{m-r}{m-1}N_{m-1}\left(r-1\right) \quad m>1.
 \end{equation}
The most commonly used form of splines is the cubic spline $\left(m=4\right)$, comprised of third degree polynomials added together at the joining points, with an explicit expression given by
\begin{align}
3!N_{4}\left(r\right)=
\begin{cases}
r^3 & \text{if $ 0\le r\le1$}, \\
4-12r+12r^2-3r^3 & \text{if $1\le r\le2$}, \\
-44+60r-24r^2+3r^3 & \text{if $2\le r\le3$}, \\
64-48r+12r^2-r^3 & \text{if $3\le r\le4$}, \\
0 & \mathrm{elsewhere}.
\end{cases}
\end{align}
The convolution and inner product of B-spline functions are required in order to formulate the multi-resolution framework described in the previous section (construction of $\boldsymbol\Lambda_{x}$, $\boldsymbol\Gamma$ and $\boldsymbol\Sigma_w$). To show how the convolution of two B-splines is calculated, \eqref{SplineConvolutionIntegral} can be rewritten as $(m-1)$ convolutions of the indicator function with itself
\begin{equation}\label{eq:N1convolutions}
 N_{m}\left(r\right)=\underbrace{\left(N_{1}\ast N_{1}\ast \cdots \ast N_{1}\right)}_{m-1\quad \text{convolutions}}\left(r\right).
\end{equation}
Using the associativity property of convolution, we have
\setlength{\arraycolsep}{0.0em}
\begin{align}\label{eq:BsplineConvolution}
N_{m}\left( r\right) \ast N_{m'}\left(r\right)&=\underbrace{\overbrace{\left(N_{1} \ast \cdots \ast N_{1}\right)}^{m-1 \quad \text{convolutions}}\left(r\right) \ast \overbrace{\left(N_{1} \ast \cdots \ast N_{1}\right)}^{m'-1\quad \text{convolutions}}}_{m+m'-1 \quad \text{convolutions}}\left(r\right)\nonumber\\
&=N_{m+m'}\left(r\right).
\end{align}
A direct consequence of \eqref{eq:BsplineConvolution} is the inner product between two B-splines is another translated B-spline such that
\begin{align}
 \left\langle N_{m}\left(r-l_{1}\right), N_{m'}\left(r-l_{2}\right)\right\rangle=&N_{m+m'}\left(m+l_{1}-l_{2}\right)\nonumber \\
=&N_{m+m'}\left(m'+l_{2}-l_{1}\right),
\label{eq:BsplineInnerProduct}
\end{align}
where $\left\langle \cdot,\cdot\right\rangle $ denotes the inner product. This holds as the support of $N_m\left(r\right)$ is $\left[ 0,m\right]$ and  is symmetric with respect to $r=\frac{m}{2}$, i.e. $ N_{m}\left(\frac{m}{2}+r\right)=N_{m}\left(\frac{m}{2}-r\right)$ (see Appendix~\ref{ap:InnerProductOfBsplines} for explicit derivation). The values of $N_{m+m'}$ in \eqref{eq:BsplineInnerProduct} can be easily determined recursively by evaluating \eqref{eq:MRA-DoBoorFormula} at integer points, i.e.
 \begin{equation}\label{eq:MRA-recursiveBsplineatintegerpoints}
 \begin{cases}
 N_2(k)=\delta(k-1)\quad k\in \mathbb{Z}, \\
 N_{m}\left(k\right)=\frac{k}{m-1}N_{m-1}\left(k\right)+\frac{m-k}{m-1}N_{m-1}\left(k-1\right) \quad k=1,2,\dots,m-1
  \end{cases}
 \end{equation}
Note $N_{m}\left(k\right)=0$ for $k\le0$ or $k\ge m$.

>>>>

The $m$-th  order cardinal B-spline function is defined by the recurrence relation \cite{Chui1992} 
\begin{equation}
N_{m}\left(r\right) = \left(N_{m-1}\ast N_{1}\right)\left(r\right) = \int_0^{1} N_{m-1}\left( r-r'\right)\,\mathrm{d}r',
\label{SplineConvolutionIntegral}
\end{equation}
where $m>1$, $\ast$ denotes convolution, and $N_1\left(r\right)$ is the characteristic function of the unit interval $\left[ 0,1\right)$, also known as indicator function, i.e.
\begin{equation}
N_{1}\left(r\right)=
\begin{cases}
1 & \text{if $ 0\le r<1$}, \\
0 & \mathrm{elsewhere}.
\end{cases}
\end{equation}
Following this, a B-spline of any order, $N_m(r)$, can be computed using the recursive expression defined by \cite{DeBoor2001}
\begin{equation}\label{eq:MRA-DoBoorFormula}
 N_{m}\left(r\right)=\frac{r}{m-1}N_{m-1}\left(r\right)+\frac{m-r}{m-1}N_{m-1}\left(r-1\right) \quad m>1.
 \end{equation}
The most commonly used form of splines is the cubic spline $\left(m=4\right)$, comprised of third degree polynomials added together at the joining points, with an explicit expression given by
\begin{align}
3!N_{4}\left(r\right)=
\begin{cases}
r^3 & \text{if $ 0\le r\le1$}, \\
4-12r+12r^2-3r^3 & \text{if $1\le r\le2$}, \\
-44+60r-24r^2+3r^3 & \text{if $2\le r\le3$}, \\
64-48r+12r^2-r^3 & \text{if $3\le r\le4$}, \\
0 & \mathrm{elsewhere}.
\end{cases}
\end{align}
\parham{One important feature of B-spline is the \emph{partition of unity} property, meaning that the unity can be expressed as a linear sum of B-splines, i.e
\begin{equation}
	\sum_{l}N_m(r+l)\equiv1
	\end{equation}
	This is important as the underlying field can be reconstructed in saturation without introducing numerical errors which is not the case for example in the case of Gaussian basis functions. }

The convolution and inner product of B-spline functions are required in order to formulate the multi-resolution framework described in the previous section (construction of $\boldsymbol\Lambda_{x}$, $\boldsymbol\Gamma$ and $\boldsymbol\Sigma_w$). To show how the convolution of two B-splines is calculated, \eqref{SplineConvolutionIntegral} can be rewritten as $(m-1)$ convolutions of the indicator function with itself
\begin{equation}\label{eq:N1convolutions}
 N_{m}\left(r\right)=\underbrace{\left(N_{1}\ast N_{1}\ast \cdots \ast N_{1}\right)}_{m-1\quad \text{convolutions}}\left(r\right).
\end{equation}
Using the associativity property of convolution, we have
\setlength{\arraycolsep}{0.0em}
\begin{align}\label{eq:BsplineConvolution}
N_{m}\left( r\right) \ast N_{m'}\left(r\right)&=\underbrace{\overbrace{\left(N_{1} \ast \cdots \ast N_{1}\right)}^{m-1 \quad \text{convolutions}}\left(r\right) \ast \overbrace{\left(N_{1} \ast \cdots \ast N_{1}\right)}^{m'-1\quad \text{convolutions}}}_{m+m'-1 \quad \text{convolutions}}\left(r\right)\nonumber\\
&=N_{m+m'}\left(r\right).
\end{align}
A direct consequence of \eqref{eq:BsplineConvolution} is the inner product between two B-splines is another translated B-spline such that
\begin{align}
 \left\langle N_{m}\left(r-l_{1}\right), N_{m'}\left(r-l_{2}\right)\right\rangle=&N_{m+m'}\left(m+l_{1}-l_{2}\right)\nonumber \\
=&N_{m+m'}\left(m'+l_{2}-l_{1}\right),
\label{eq:BsplineInnerProduct}
\end{align}
where $\left\langle \cdot,\cdot\right\rangle $ denotes the inner product. This holds as the support of $N_m\left(r\right)$ is $\left[ 0,m\right]$ and  is symmetric with respect to $r=\frac{m}{2}$, i.e. $ N_{m}\left(\frac{m}{2}+r\right)=N_{m}\left(\frac{m}{2}-r\right)$ (see Appendix~\ref{ap:InnerProductOfBsplines} for explicit derivation). The values of $N_{m+m'}$ in \eqref{eq:BsplineInnerProduct} can be easily determined recursively by evaluating \eqref{eq:MRA-DoBoorFormula} at integer points, i.e.
 \begin{equation}\label{eq:MRA-recursiveBsplineatintegerpoints}
 \begin{cases}
 N_2(k)=\delta(k-1)\quad k\in \mathbb{Z}, \\
 N_{m}\left(k\right)=\frac{k}{m-1}N_{m-1}\left(k\right)+\frac{m-k}{m-1}N_{m-1}\left(k-1\right) \quad k=1,2,\dots,m-1
  \end{cases}
 \end{equation}
Note $N_{m}\left(k\right)=0$ for $k\le0$ or $k\ge m$.
